Simulation Tools for Small Area Estimation: Introducing the \proglang{R}-package \proglang{saeSim}

### 2014-10-28 saeSim.tex
Update saeSim paper

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
%
\section{Introduction}
The demand for reliable small area statistics from sample surveys has been substantial grown over the last decades due to their use in public and private sectors. In this paper we present a framework for simulation studies inside the field of small area estimation. This tool might be useful for the prospective researcher or data analyst to provide reproducible research.

Reproducible research has become a widely discussed topic in many academic areas. In the field of statistics many mostly open-source tools like the \proglang{R}-language \citep{r14} and \LaTeX, dynamic reporting packages like \proglang{knitr} \citep{yihui13}, \proglang{sweave} \citep{leisch02} and more recently \proglang{rmarkdown} \citep{allaire14}, make the integration of text and source code for statistical analysis possible. Publishing source code and data alongside research results draws special attention to authoring the analysis. However, the requirements for source code are different from the written words in the article itself.

@@

\subsection{Model-based simulation}
\label{sec:csModel}
In the following we want to show how to construct a simulation in a model-based setting. The goal is to estimate the domain predictions under a FH model. Components which are involved are \textit{data generation} and \textit{computing on aggregated data}. The first step is to generate the data under the model:

@@

\begin{Schunk}
\begin{Sinput}
> library(saeSim)
\end{Sinput}
\end{Schunk}
\begin{Schunk}
\begin{Sinput}
> setup <- base_id(nDomains = 40, nUnits = 1) %>% 
+   sim_gen_x(mean = 0, sd = 4) %>%
+   sim_gen_v(mean = 0, sd = 1)
> setup
\end{Sinput}
\begin{Soutput}
  idD          x          v
1   1 -2.5058152 -0.1645236
2   2  0.7345733 -0.2533617
3   3 -3.3425144  0.6969634
4   4  6.3811232  0.5566632
5   5  1.3180311 -0.6887557
6   6 -3.2818735 -0.7074952
\end{Soutput}
\end{Schunk}

Note that if you print a simulation setup to the console, as in the above example, one simulation run is performed and only the first rows of the resulting data table are printed. This enables interactivity with the object itself, however it hides that the setup object is a collection of functions to be called. The error component $e_i$ in the model has different variances which is not covered by a predefined function. Thus, as a \textit{generator component} we define a function which takes a \proglang{data.frame} as input and returns it after adding a variable named \proglang{vardir} with the variances and the variable \proglang{e} with the generated random numbers:

@@

To add the area-level predictions from a Fay-Herriot model we need to define another component. The function takes a \proglang{data.frame} as input and returns the modified version. For the estimation of the EBLUP under the FH model we use the function \proglang{eblupFH} from the package \proglang{sae} \citep{molina13}. Finally we add \proglang{comp\_FH} to the process:

\begin{Schunk}
\begin{Sinput}
> library(sae)
\end{Sinput}
\end{Schunk}
\begin{Schunk}
\begin{Sinput}
> comp_FH <- function(dat) {
+   modelFH <- eblupFH(y ~ x, vardir, data = dat)
+   dat$FH <- modelFH$eblup
+   dat
+ }
> setup <- setup %>% sim_comp_agg(comp_FH)
> setup
\end{Sinput}
\begin{Soutput}
  idD         x          v vardir          e         y  trueStat        FH
1   1  1.637607  0.7073107    0.1  0.1258998 104.10843 103.98253 104.06121
2   2  6.755493  1.0341077    0.2 -0.1822523 114.36284 114.54509 114.23876
3   3  6.346354  0.2234804    0.3  0.7253263 113.64151 112.91619 113.45213
4   4 -1.323631 -0.8787076    0.4 -0.4434978  96.03053  96.47403  96.45416
5   5 -9.140942  1.1629646    0.5 -0.4105563  82.47052  82.88108  82.46045
6   6  9.990646 -2.0001649    0.6 -0.7754272 117.20570 117.98113 118.08379
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> library(dplyr)
\end{Sinput}
\end{Schunk}
\begin{Schunk}
\begin{Sinput}
> simResults <- sim(setup, R = 100) %>% rbind_all
> simResults %>% select(idD, idR, simName, trueStat, y, FH)
\end{Sinput}
\begin{Soutput}
Source: local data frame [4,000 x 6]

@@

An additional variable \proglang{idR} is automatically added as an ID-variable for the iteration as well as a variable \proglang{simName} to distinguish between scenarios, should there be more than one. At this time we do not provide further tools to process the resulting data. As it is a \proglang{data.frame} many packages are available in \proglang{R}. This point will be presented in more detail in the design-based study.

@@

\begin{Schunk}
\begin{Sinput}
> library(ggplot2)
> ggplot(ggDat, aes(x = region, y = prediction - trueStat, fill = method)) + 
+   geom_boxplot() + theme(legend.position = "bottom")
\end{Sinput}
\end{Schunk}

\includegraphics[width = \textwidth]{saeSim-designSimBIAS.pdf}

\begin{Schunk}
\begin{Sinput}
> ggplot(ggDat, aes(x = region, y = (prediction - trueStat)^2, fill = method)) + 
+   geom_boxplot() + scale_y_log10() + theme(legend.position = "bottom")
\end{Sinput}
\end{Schunk}

\includegraphics[width = \textwidth]{saeSim-designSimMSE.pdf}

>>>>

\begin{document}
%
%% include your article here, just as usual
%
% R CMD Sweave %.Rnw
% Loading (and installing) necessary packages.
%
\section{Introduction}
The demand for reliable small area statistics from sample surveys has been substantially grown over the last decades due to their use in public and private sectors. In this paper we present a framework for simulation studies inside the field of small area estimation. This tool might be useful for the prospective researcher or data analyst to provide reproducible research.

Reproducible research has become a widely discussed topic. In the field of statistics many mostly open-source tools like the \proglang{R}-language \citep{r14} and \LaTeX, dynamic reporting packages like \proglang{knitr} \citep{yihui13}, \proglang{sweave} \citep{leisch02} and more recently \proglang{rmarkdown} \citep{allaire14}, make the integration of text and source code for statistical analysis possible. Publishing source code and data alongside research results draws special attention to authoring the analysis. However, the requirements for source code are different from the written words in the article itself.

@@

\subsection{Model-based simulation}
\label{sec:csModel}
In the following we want to show how to construct a simulation in a model-based setting. The goal is to estimate the domain predictions under a FH model. Components which are involved are \textit{data generation} and \textit{computing on aggregated data} (see figure \ref{fig:flowdiagram}). The first step is to generate the data under the model:

@@

\begin{Schunk}
\begin{Sinput}
> library(saeSim)
> setup <- base_id(nDomains = 40, nUnits = 1) %>% 
+   sim_gen_x(mean = 0, sd = 4) %>%
+   sim_gen_v(mean = 0, sd = 1)
> setup
\end{Sinput}
\begin{Soutput}
  idD          x          v
1   1 -2.5058152 -0.1645236
2   2  0.7345733 -0.2533617
3   3 -3.3425144  0.6969634
4   4  6.3811232  0.5566632
5   5  1.3180311 -0.6887557
6   6 -3.2818735 -0.7074952
\end{Soutput}
\end{Schunk}

Note that if you print a simulation setup to the console, as in the above example, one simulation run is performed and only the first rows (the head) of the resulting data table are printed. This enables interactivity with the object itself, however it hides that the setup object is a collection of functions to be called. In this model the error component $e_i$ has different variances which is not covered by a predefined function. Thus, as a \textit{generator component} we define a function which takes a \proglang{data.frame} as input and returns it after adding a variable named \proglang{vardir} with the variances and the variable \proglang{e} with the generated random numbers:

@@

To add the area-level predictions from a Fay-Herriot model we need to define another component. The function takes a \proglang{data.frame} as input and returns the modified version. For the estimation of the EBLUP under the FH model we use the function \proglang{eblupFH} from the package \proglang{sae} \citep{molina13}. Hence we define a function named \proglang{comp\_FH} and add it to the process:

\begin{Schunk}
\begin{Sinput}
> library(sae)
> comp_FH <- function(dat) {
+   modelFH <- eblupFH(y ~ x, vardir, data = dat)
+   dat$FH <- modelFH$eblup
+   dat
+ }
> setup <- setup %>% sim_comp_agg(comp_FH)
> setup
\end{Sinput}
\begin{Soutput}
  idD         x          v vardir          e         y  trueStat        FH
1   1  1.637607  0.7073107    0.1  0.1258998 104.10843 103.98253 104.06121
2   2  6.755493  1.0341077    0.2 -0.1822523 114.36284 114.54509 114.23876
3   3  6.346354  0.2234804    0.3  0.7253263 113.64151 112.91619 113.45213
4   4 -1.323631 -0.8787076    0.4 -0.4434978  96.03053  96.47403  96.45416
5   5 -9.140942  1.1629646    0.5 -0.4105563  82.47052  82.88108  82.46045
6   6  9.990646 -2.0001649    0.6 -0.7754272 117.20570 117.98113 118.08379
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
> library(dplyr)
> simResults <- sim(setup, R = 100) %>% rbind_all
> simResults %>% select(idD, idR, simName, trueStat, y, FH)
\end{Sinput}
\begin{Soutput}
Source: local data frame [4,000 x 6]

@@

An additional variable \proglang{idR} is automatically added as an ID-variable for the iteration as well as a variable \proglang{simName} to distinguish between scenarios, should there be more than one. At this time we do not provide further tools to process the resulting data. As it is a \proglang{data.frame} many packages are available in \proglang{R}. In the design-based scenario we show how to process the result data into graphs with only a few lines of code.

@@

\begin{Schunk}
\begin{Sinput}
> library(ggplot2)
> ggplot(ggDat, aes(x = region, y = prediction - trueStat, fill = method)) + 
+   geom_boxplot() + theme(legend.position = "bottom")
> ggplot(ggDat, aes(x = region, y = (prediction - trueStat)^2, fill = method)) + 
+   geom_boxplot() + scale_y_log10() + theme(legend.position = "bottom")
\end{Sinput}
\end{Schunk}

\begin{figure}[!h]
\includegraphics[width = \textwidth]{saeSim-designSimBIAS.pdf}
\caption{Monte-Carlo BIAS of direct vs. BHF predictor. 50 predictions for each region and estimation technique.}
\label{fig:dbBIAS}
\end{figure}

\begin{figure}[!h]
\includegraphics[width = \textwidth]{saeSim-designSimMSE.pdf}
\caption{Monte Carlo MSE of direct vs. BHF predictor. 50 predictions for each region and estimation technique.}
\label{fig:dbMSE}
\end{figure}
